---
title: "COVID School Closures and Their Impact on Students"
<<<<<<< HEAD
author: "Cristina Burca, Yan Mezhiborsky, Sakura Noskor"
=======
author: "Cristina Burca, Yan B. Mezhiborsky, Sakura"
>>>>>>> cc5671c5ca1423620234eea34918dd30fa422491
format: pdf
date: 02/12/2024
date-format: long
thanks: "Code and some data from this paper are available at: [github repo](https://github.com/NotSakura/COVIDSchoolClosures.git)"
classoption: abstract
abstract: |
 COVID-19 was an outbreak of a virus that forced many institutions to shut down for 2-3 years. Schools were no different and this paper aims to look at the effects of the said closures in school and how it affected the students. This paper's initial focus is on the global student population, where it finds that there were multiple days of school closed during covid, but shifts its focus to the US where it finds that the standardized test scores have also been affected. Other results was that with more in-person schooling provided the less the enrollment rates dropped (more in-depth analysis in later sections). 
include-in-header: 
  text:
    \renewcommand{\abstractname}{Abstract}
bibliography: references.bib
margins:
  top: 0.5in
  bottom: 0.5in
  left: 0.5in
  right: 0.5in
toc: true
---

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

library(knitr)
library(janitor)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(haven)
library(tsibble)
library(ggplot2)
library(countrycode)
library(lubridate)
```

# Introduction

This paper will discuss several ways that the Covid-19 pandemic may have affected schools students and education as a whole. We will be looking at data procuded from all around the world , although we will be focusing on the data collected from the United States of America. First, we will be looking at school closures in the five major continents and some dductions we can make from them. Next, we will look at test scores for the USA in Mathematics as well as ELA(English Learning Arts), and how they changed from year to year, before, during, and after the pandemic. Lastly, the outcomes of this paper are that there is an over all increase in the percent of enrollment as more in person schools are opened in the educational districts of United States. The concept fo educational districts will be described later in the data section for the user's aid but in this context it is not the most important.

We will also discuss the differences in our findings from the original paper as well as addition analasis from the data provided from their replication package.


# Data

## Data Source and Collection

We use @citeR to make this paper as well as the graphs and topic were taken inspiration from @CovidPaper. Various helpful packages were used in order to clean, sort and graph this paper in a way such that the reader will not have difficulty undertsnading neither the topic nor the data sets of this paper. The packages are, @Ggplot2, @tidyverse, @rDplyr, @rReadr, @rknitr, @rJanitor, @rExtra, @rTidyr, @rHaven, @rTissible.

## Data Cleaning / Methodology

What data set did we clean and why. Explain the variable here too ig The data provided originally was called nces Some data sets cleaned were


  The third set of graphs shown in this paper is the enrollment rates in the different educational districts of the US.The original data provided was from US department of Education through NCES (National Institute for Education Statistics). The variables included were year, leaid, fips, grade, race, sex, and enrollment. Starting by defining the variables, leaid stands for local education agency id number. This allows the NCES to identify the districts with their unique id code. Next is fips, which stands for federal information processing standards. This number is a rate/level given by the US government to a specific district which essentially rates the technological security of each district (@fipsDef). The grade essentially says the grade the students are in from 0-12 to 99 where 0 is kindergarten and 99 is representing the total enrollment for all grades. The dataset was also made through another raw data set from COVID School Data Hub [@CSDH] which contains different abbreviations and id numbers corresponding to each district, including leaid, and they also have the share of in-person, virtual and hybrid schools open in that district. The share of in-person school represents the amount of schools that were in_person in that district. Similar explanations for virtual and hybrid shares. 

  The process of cleaning this raw data was simple. First, we look at the NCES data. Since the paper only looks at enrollment rates for all grades and kindergartners, a filter is applied for grade to be 0 or 99 with another filter of race to be 99 as 99 represents all. Then another filter is applied where, the data only contains fips that has a value less than 56 as there are only 56 levels in fips.  Then the entire data is transformed into a wide format where the new following new columns appear: enrollment for males, enrollment for females, enrollment for all (Not the real names). There are other columns like enrollment9 that gets created but it is dropped as there are no values contained within it. Then the values of enrollment for all columns are filtered such that if they contain an invalid value or such as not available or 0, then the entire row is deleted. After that, the entire data is again reshaped such that we get rid of the grades columns and have new columns as such; enrollment for males in total, enrollment for females in total, enrollment for males in kindergarten, enrollment for females in kindergarten, enrollment for both gender in total, enrollment for both gender in kindergarten. So currently have the rows listed before, the year, and the leaid. Then the current data is merged with the other raw data provided by COVID School Data Hub[@CSDH]. They were merged based on the leaid and any access columns such as NCES id or state id was deleted as it was of no use to the paper. Then the computations begin. In order to find the rate of enrollment, the formula was to subtract the previous value from the current and divide by the previous value. So for example to find the enrollment rate for males that attend in-person, the `lag()` of enrollment for males in total was taken and subtracted to it self and stored in a variable called changeTotal. Then the changeTotal variable was divided again by the lag of "enrollment for males in total" column. This was repeated for all the columns that had the enrollment values for each subcategory. This value was then binned according to the share in-person column or the share virtual column based on the graph that was produced. However this was not enough as there were many points, to the point where the analysis of that data set was not worth anything and hence, per guidance to the original paper, a weighted mean was taken. An ordinary average was not taken as this some states would have higher in-person rates then others which would mean there are outliars in this case. So an ordinary average function would have lost some of the important trends we see when we take a weighted average. 

## Measurment
The second set of this paper's graphs focused on test scores of students in the United States of America, specifically their score changes in ELA(English Language Arts) and Mathematics. The data this was compared to was the percentage of in person attendance as well as the geographic locale of the students.

As mentioned before the third claim that this paper looks at is the enrollment rates with respect to the share of in-person learning or virtual learning. For this category there are 2 measurements. The data originally come with the number of enrollments in each leaid for each year however, the data is cleaned up so the reader may see the rates related to it. The x-axis show the share of in-person or virtual schools in the district. The share of in-person was only changed in terms of how this paper binned the data; the share value was from 0 to 100% and it was cut into 11 equal pieces. The measurement for the x-axis is the same as it is on the raw datasheet, which is percentage representing share of districts that were either in-person or virtual, values from 0 to 100%. The measurement for the y axis is the enrollment rate. This has been changed from the raw data set as originally it recorded the number of students enrolled for each grade by leaid for a certain year. As mentioned in the methodology, the original value was changed into a rate then a percentage to better show the reader what the actual changes were. If this step was not done then comparing the declines and the inclines of enrollment rate would have been hard due to the large numbers. Anyhow, the measurement for the y-axis is the enrollment rate in percentage and it doesn't have a limit. 




# Results


## First Graph

```{r}
#| echo: false
#| include: false

covid <- read.csv("../../inputs/data/OxCGRT_compact_national_v1.csv")
#cleaning dates column
##use as.date for expected date format, as.character to convert column to string
covid$Date <- as.Date(as.character(covid$Date),
                      format="%Y%m%d")

#create new column for years
covid$Year <- year(covid$Date)

#filtering signifcant variables
covid <- covid %>% 
  select(Date, Year, CountryName, C1M_School.closing)

#using package 'countrycode', create new column 'continent' that lists continent of country
##sourcevar is input, origin tells function input is list of countries
##origin
covid$Continent <- countrycode(sourcevar = covid$CountryName,
                            origin = "country.name",
                            destination = "continent")

covid <- filter(covid, !is.na(Continent))
covid<- covid %>% group_by(Year)
```

As the original paper has divided the countries by Region, this paper has divided the countries by continent, to compare and contrast the findings based on the separation of the countries.

The original paper states that in their findings, the United States had longer terms of school closures, while given by the graph I have produced, it seems

```{r}
#| echo: false
#| warning: false

#graph 2

covid_summary <- covid %>%
  group_by(Year, Continent) %>%
  summarise(Total_School_Closings = sum(C1M_School.closing, na.rm = TRUE))

#reordering in descending order
covid_summary$Continent <- reorder(covid_summary$Continent, -covid_summary$Total_School_Closings)

#
ggplot(covid_summary, aes(x = Continent, y = Total_School_Closings, fill = Continent)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Continent", y = "Total School Closings",
       title = "Total School Closings by Continent")

ggplot(covid_summary, aes(x = Continent, y = Total_School_Closings, fill = Continent)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  facet_wrap(~Year) +
  labs(x = "Continent", y = "Total School Closings",
       title = "Total School Closings by Continent")
```

## Second Graph

Here we will look at the change in percentage of test score of students between the years of 2017 and 2022. Covid-19 had a major impact on student test scores, whether it be because of the quality of remote learning or the resources the students may have or not have had while learning remotely. the first of the two figures is a near identical replication of the the on in the original paper we are replicating while the second figure is gives us some additional information to analyze that wasn't examined as thoroughly in the original paper.


<<<<<<< HEAD
```{r, fig.align='center'}
=======
```{r}
#| include: false
>>>>>>> cc5671c5ca1423620234eea34918dd30fa422491
#| echo: false
#| warning: false
#| label: fig-InPersonAttendence
#| fig-cap: Average Grade Change by In Person Attendence

score_data = read_csv("../../inputs/data/scores_lm_demographics.csv")


clean_score_data_inperson <- score_data |>
  select(subject, change_2017_2018, change_2018_2019, change_2019_2021, change_2021_2022, share_inperson, urban_centric_locale) |>
  rename_with(~ sub("^change_(\\d{4})_(\\d{4})$", "Spring_\\2", .), starts_with("change")) |>
  mutate(share_inperson_grouped = cut(share_inperson * 100, breaks = seq(0, 100, by = 25), include.lowest = TRUE, right = FALSE, labels = c("0-25", "25-50", "50-75", "75-100")))

# Pivot the data to a long format
score_data_long_inperson <- clean_score_data_inperson |>
  pivot_longer(cols = starts_with("Spring"), names_to = "time_period", values_to = "change_score")

# Group by 'subject', 'share_inperson_grouped', and 'time_period', then summarize
score_data_summary_inperson <- score_data_long_inperson |>
  group_by(subject, share_inperson_grouped, time_period) |>
  summarise(
    mean_change = mean(change_score, na.rm = TRUE),
    .groups = 'drop'
  )

# Now prepare data for the 'urban_centric_locale' grouping
score_data_long_locale <- clean_score_data_inperson |>
  pivot_longer(cols = starts_with("Spring"), names_to = "time_period", values_to = "change_score")

# Group by 'subject', 'urban_centric_locale', and 'time_period', then summarize
score_data_summary_locale <- score_data_long_locale |>
  group_by(subject, urban_centric_locale, time_period) |>
  summarise(
    mean_change = mean(change_score, na.rm = TRUE),
    .groups = 'drop'
  )

ggplot(score_data_summary_inperson, aes(y = share_inperson_grouped, x = round(mean_change * 100), color = time_period)) +
  geom_point(position = position_dodge(width = 0.2)) +
  scale_x_continuous(limits = c(-15, 10), breaks = seq(-15, 10, by = 5)) +
  labs(
    title = "Average Grade Change by In Person Attendence",
    y = "In-Person Share Group (%)",
    x = "Average Change Score (%)",
    color = "Time Period" 
  ) +
  scale_color_brewer(palette = "Set1", labels = c("Spring 2018", "Spring 2019", "Spring 2021", "Spring 2022")) + # Custom labels
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    legend.background = element_rect(fill = "white", size = 0.3, linetype = "solid"),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10, face = "bold"),
    legend.key.size = unit(0.2, "cm")
  ) +
  facet_wrap(~subject)


```

The first observation that can be made is the percent change in test score between Spring 2019 and Spring 2021 where the students who who attended class less than 25% of the time had an average of 15% less in their Math tests and approximately 9% in their ELA tests. Although it can be seen that the more the students attended in person the less their test scores would decrease, never the less the students still saw a decrease in their test scores between the two years even if they had gone to school in person. One thing the original paper failed to include although it was in their data set, was the change in test scores of Spring 2022. While the test scores do increase in Math and ELA, ELA less so than Math, they do not increase as much as they fell in the previous year.



<<<<<<< HEAD
```{r, fig.align='center'}
=======
```{r}
#| include: false
>>>>>>> cc5671c5ca1423620234eea34918dd30fa422491
#| echo: false
#| warning: false
#| label: fig-GeoLocale
#| fig-cap: Average Grade Change by In Person Attendence


# Plot for 'urban_centric_locale'
ggplot(score_data_summary_locale, aes(y = urban_centric_locale, x = round(mean_change * 100), color = time_period)) +
  geom_point(position = position_dodge(width = 0.1)) +
  scale_x_continuous(limits = c(-15, 10), breaks = seq(-15, 10, by = 5)) +
  labs(
    title = "Average Score Change by Geographic Locale",
    y = "Urban Centric Locale",
    x = "Average Change Score (%)",
    color = "Time Period" 
  ) +
  scale_color_brewer(palette = "Set1", labels = c("Spring 2018", "Spring 2019", "Spring 2021", "Spring 2022")) + # Custom labels for time periods
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", size = 0.5, linetype = "solid"),
    legend.text = element_text(size = 8), 
    legend.title = element_text(size = 10, face = "bold"), 
    legend.key.size = unit(0.2, "cm")
  ) +
  facet_wrap(~subject)
```



Here we will look at the Geographic Locales of students, and how their grade have changed based on what type of municipality they live in. We can observe that Cities and suburbs of all sizes both in Math and ELA Scores. This may be due to a variety of reasons among them the Covid-19 in each of the specific localities. As cities are densely populated they are more likely to have had stricter lock down rules during the outbreak of Covid-19. Furthermore, suburbs are usually near cities and would likely have similar lock down rules to the cities they are near.

## Enrollment rate in the US
It is no secret that COVID-19 affected the school enrollment rates, when the students were forced to learn in a virtual environment. In this section the paper will uncover any trends in enrollment rates with relation to the number of in-person learning offered in each district. For all of the graphs below the y axis will be from -0.12 to 0.05 to ensure reader us able to comapre the trends right beside each other.

```{r}
#| echo: false
#| warning: false
## Only run once cause super big data which will slow down
raw_d <- read_dta("../../inputs/data/nces_district_enrollment_2018_2020.dta")
other_raw <- read_csv("../../inputs/data/District_Overall_Shares.csv")
other_raw <- other_raw|>
  rename(leaid = "NCESDistrictID")
other_raw$leaid <- as.character(other_raw$leaid)
```

```{r}
#| echo: false
#| warning: false
data <- raw_d |>
  filter(grade == 0 | grade == 99) |>
  filter(race == 99)

data <- subset(data, select = -race)

# Drop observations where enrollment is less than 0
data <- data[data$enrollment >= 0, ]

# Drop observations where fips is greater than 56
data <- data[data$fips <= 56, ]

data <- data |>
  filter(!is.na(sex))


# Reshape 'enrollment' wide by 'leaid', 'year', and 'grade'
data_wide <- pivot_wider(data,
                            names_from = sex,
                            values_from = enrollment) |>
  rename(
    enrollment1 = `1`,
    enrollment2 = `2`,
    enrollment9 = `9`,
    enrollment99 = `99`)

# Drop the 'enrollment9' variable
data_wide$enrollment9 <- NULL

# Drop observations where 'enrollment99' is missing or less than or equal to 0
data_wide <- data_wide[!is.na(data_wide$enrollment99) & data_wide$enrollment99 > 0, ]

# Reshape 'enrollment1', 'enrollment2', and 'enrollment99' wide by 'leaid' and 'year'
data_final <- pivot_wider(data_wide,
                            names_from = grade,
                            values_from = c(enrollment1, enrollment2, enrollment99))
# Can get rid of fips

data_final <- subset(data_final, select = -fips)

# Rename columns
colnames(data_final) <- c("year", "leaid", 
                              "enroll_male_total", "enroll_male_kinder", 
                              "enroll_female_total", "enroll_female_kinder", 
                              "enroll_all_total", "enroll_all_kinder")


# Convert 'leaid' to numeric (assuming it's a string)
data_final$leaid <- as.numeric(data_final$leaid)
other_raw$leaid <- as.numeric(other_raw$leaid)


write_csv(x = data_final,
          file = "../data/cleaned_raw_data.csv")


# Link data with another dataset (replace 'other_data' with your actual dataset)
# Assuming 'leaid' is the common identifier
merged_data <- merge(data_final, other_raw, by = "leaid")
merged_data <- subset(merged_data, select = -DistrictName)
merged_data <- subset(merged_data, select = -StateAbbrev)
merged_data <- subset(merged_data, select = -StateAssignedDistrictID)

write_csv(x = merged_data,
          file = "../data/merged_data.csv")


# Assuming 'year' is numeric (e.g., 2019, 2020)
merged_data$date <- as.Date(paste0(merged_data$year, "-01-01"))


my_tsibble <- as_tsibble(merged_data, key = leaid, index = date)


# Calculate change_kinder
my_tsibble <- my_tsibble |>
  mutate(change_kinder = enroll_all_kinder - lag(enroll_all_kinder),
         perc_change_kinder = change_kinder / lag(enroll_all_kinder))



my_tsibble <- my_tsibble |>
  arrange(share_inperson) |>
  mutate(change_all = enroll_all_total - lag(enroll_all_total),
         perc_change_all = change_all / lag(enroll_all_total))

# Calculate change_male
my_tsibble <- my_tsibble |>
  mutate(change_male = enroll_male_kinder - lag(enroll_male_kinder),
         perc_change_male = change_male / lag(enroll_male_kinder))

# Calculate change_female
my_tsibble <- my_tsibble |>
  mutate(change_female = enroll_female_kinder - lag(enroll_female_kinder),
         perc_change_female = change_female / lag(enroll_female_kinder))


# Drop rows where absolute change_all is greater than enroll_all_total
my_tsibble <- my_tsibble |>
  filter(abs(change_all) <= enroll_all_total)

# Drop rows where absolute change_kinder is greater than enroll_all_total
my_tsibble <- my_tsibble |>
  filter(abs(change_kinder) <= enroll_all_kinder)


virtual_tsibble <- my_tsibble |>
  mutate(virtual_bin = cut(share_virtual, breaks = 11))
my_tsibble <- my_tsibble |>
  mutate(inperson_bin = cut(share_inperson, breaks = 11))


# get the different enrollment rates for different years
enrollment_rate_2019 <- my_tsibble |>
  filter(year >= 2019)
my_tsibble <- my_tsibble |>
  filter(year >= 2020)
virtual_tsibble <- virtual_tsibble |>
  filter(year >= 2020)


# get the mean enrollment rate for 2019
mean_enroll_2019_all <- enrollment_rate_2019 |>
  group_by(inperson_bin) |>
  summarise(
    weighted_mean_change = weighted.mean(perc_change_all, enroll_all_total)
  )

mean_enroll_2019_all <- subset(mean_enroll_2019_all, select = -date)
mean_enroll_2019_all <- mean_enroll_2019_all |>
  group_by(inperson_bin) |>
  summarise_at(vars(weighted_mean_change), list(weighted_mean_change = mean)
  )

# get the mean enrollment rate for 2020 but virtual this time
virtual_2020_mean <- virtual_tsibble |>
  group_by(virtual_bin) |>
  summarise(
    weighted_mean_change = weighted.mean(perc_change_all, enroll_all_total)
  )
virtual_2020_mean <- subset(virtual_2020_mean, select = -date)


# Compute weighted mean change in enrollment for each bin
weighted_means_all <- my_tsibble |>
  group_by(inperson_bin) |>
  summarise(
    weighted_mean_change = weighted.mean(perc_change_all, enroll_all_total)
  )

weighted_means_all <- subset(weighted_means_all, select = -date)
weighted_means_all <- weighted_means_all |>
  group_by(inperson_bin) |>
  summarise_at(vars(weighted_mean_change), list(weighted_mean_change = mean)
  )


weighted_means_kinder <- my_tsibble |>
  group_by(inperson_bin) |>
  summarise(
    weighted_mean_change = weighted.mean(perc_change_kinder, enroll_all_kinder)
  )
weighted_means_kinder <- subset(weighted_means_kinder, select = -date)

```



```{r}
#| echo: false
#| warning: false
#| label: fig-EnrollAll
#| fig-cap: The weighted mean of the percent of overall district enrollment was calulated and graphed. The weights were decided on the bins which were cut in eleven sections based on share of in person school overall. 


##Graph
# Extract the end value and convert to numeric
weighted_means_all$inperson_bin <- as.numeric(sub(".*,", "", sub("\\]", "", weighted_means_all$inperson_bin)))



g <- ggplot(weighted_means_all, aes(x = inperson_bin , y = weighted_mean_change)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE)+
  scale_y_continuous(breaks = seq(-0.12, 0.05, by = 0.02), limits = c(-0.12, 0.05))+
  labs(
    title = "Percent Change in Overall District Enrollment, Fall 2019-Fall 2020",
    x = "Share of School Year with In-Person Learning Offered, 2020-2021",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1, size = 6))

ggsave(filename = "../figures/InpersonAll.png", plot = g, width = 6, height = 4)
g

```

First, we take a look at the overall enrollment through out the districts. We can see that most values are in the negatives with one value in the positives in the \[0.545, 0.6360\] bin. We however see an upward trend signifying that we have a a positive change in the percent of rate of change. This means that the enrollment rate increases with more shares of the district going in person. Diving in deeper to look at the values (provided in the Appendix).The graph shows that for 0% to 9% of in-person schooling, we have a decrease of 3% in the enrollment rate. For 9% to 55% share or in person schooling, the rate of enrollment is around -35% meaning there is  decrease of 35% and the rate goes to 4% increase in enrollment rate for 55% to 64% of in person schooling in the district. Then the value drops to 1% decrease in enrollment rate when the share of in person offered is 73% to 64% approximately. Then it jumps to a 2% decrease in enrollment then it levels out to 1% decrease in enrollment rate for the last 73% to 100% share in in-person learning. The overall trend is increasing but there are brief moments of decreasing enrollment rates. 

```{r}
#| echo: false
#| warning: false
#| label: fig-EnrollKinder
#| fig-cap: The weighted mean of the percent of kindergarden enrollment in all districts was calulated and graphed. The weights were decided on the bins which were cut in eleven sections based on share of in person school overall.  

weighted_means_kinder$inperson_bin <- as.numeric(sub(".*,", "", sub("\\]", "", weighted_means_all$inperson_bin)))


g <- ggplot(weighted_means_kinder, aes(x = inperson_bin, y = weighted_mean_change)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE)+
  scale_y_continuous(breaks = seq(-0.12, 0.05, by = 0.02), limits = c(-0.12, 0.05))+
  labs(
    title = "Percent Change in Kindergarten Enrollment, Fall 2019-Fall 2020",
    x = "Share of School Year with In-Person Learning Offered, 2020-2021",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1, size = 6))

ggsave(filename = "../figures/InpersonKinder.png", plot = g, width = 6, height = 4)
g
```
In comparison we look at the kindergarten enrollment in the same year. Looking at the values individually, the graph shows that for the initial shares 9%, 18% and 27%, we get an enrollment rate of kindergartners in the US is -9%, -10% and 10% respectively. They all show decreasing enrollment rate. The for the 36% of in-person schooling we get a 8.4% enrollment rate and there is a decrease of 9.4% and 8.2% in enrollment rate when there is 45% and 55% share of in person school respectively, for kindergartens. Then the enrollment rates increase to -3.2% and 7% for the proceeding 64% and 73% of in person school share. Afterwards there is a decrease 4.8% in enrollment rates when 82% of schools are in person. Then there is another decrease to -2.4% in enrollment rate when there is 91% share of in person rates. And lastly for 91% to 100% of the in person shares the enrollment rates are -3.9%. This means that although the enrollment rates are negative for the most part there is an increasing trend. This means that the enrollment rates were increasing as more kindergartners were attending in person. Also compared to @fig-EnrollAll, the slope of the line of best fit is steeper meaning that the rate of increas in enrollment rates is higher. More on the reason to be discussed in later sections. 

```{r}
#| echo: false
#| warning: false
#| label: fig-Enroll2019
#| fig-cap: The weighted mean of the percent of total enrollment in 2019 accross all districts was calulated and graphed. The weights were decided on the bins which were cut in eleven sections based on share of in person school overall.  

mean_enroll_2019_all$inperson_bin <- as.numeric(sub(".*,", "", sub("\\]", "", mean_enroll_2019_all$inperson_bin)))


g <- ggplot(mean_enroll_2019_all, aes(x = inperson_bin, y = weighted_mean_change)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE)+
  scale_y_continuous(breaks = seq(-0.12, 0.05, by = 0.02), limits = c(-0.12, 0.05))+
  labs(
    title = "Percent Change in Overall Enrollment, Fall 2018-Fall 2019",
    x = "Share of School Year with In-Person Learning Offered, 2019-2020",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1, size = 6))

ggsave(filename = "../figures/InpersonAll2019.png", plot = g, width = 6, height = 4)
g
```

The purpose of this paper is to compare and analyse the outcomes of the school closures during the pandemic to better understand the impact of it. To see the comparison, this paper also analyses the enrollment rate in 2019. The graph shows that the initial 9% share of in-person school has a decrease of 1% in enrollment.These values stay relatively the same - hovering around the -0.10% to -0.019% enrollment rate- until the 55% to 64% share of in-person school where the enrollment rate jumps to 0.02. Then in the 73th percentile, we see a decrease to -1% in enrollment rate but the procceeding values even out to approximatly 0$ (or more specifically 0.001% and 0.004%). Although the values are negative for the large portion of the graph, there is still an upward trend we saw in previous graphs where the more proprtion of in-person schooling, the larger the enrollment rate. 

```{r}
#| echo: false
#| warning: false
#| label: fig-EnrollVirtual
#| fig-cap: The weighted mean of the percent of total enrollment in 2020 accross all districts was calulated and graphed. The weights were decided on the bins which were cut in eleven sections based on share of virtual school overall.  

virtual_2020_mean$virtual_bin <- as.numeric(sub(".*,", "", sub("\\]", "", virtual_2020_mean$virtual_bin)))


g <- ggplot(virtual_2020_mean, aes(x = virtual_bin, y = weighted_mean_change)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE)+
  scale_y_continuous(breaks = seq(-0.12, 0.05, by = 0.02), limits = c(-0.12, 0.05))+
  labs(
    title = "Percent Change in Overall Enrollment, Fall 2019-Fall 2020",
    x = "Share of School Year with Virtual Learning Offered, 2020-2021",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1, size = 6))

ggsave(filename = "../figures/VirtualAll.png", plot = g, width = 6, height = 4)
g
```

Lastly, in order to properly make a conclusion about enrollment rates and in-person schooling the reader must also see the opposite correlation, hence the paper looks at the enrollment rates for share of virtual learning as well, above. The graphs shows that for the first 9% of schools doing virtual learning in the districts have a enrollment rate of -2%, which increase to approximately -3% in the 18% share of virtual school. Then the value jumps up to the maximum value in the entire graph at 0.001% decrease in enrollment rate when 18% to 27% of schools are virtual. Then for the shares from 27% to 72% we see that the enrollment rate fluctuates between 3.1 to 3.7% decrease in enrollment rate where the change is so small the difference is not changing the graph at all. Finally we see a slight increase in enrollment rate with -2.8% and then proceeding with an even higher enrollment rate of -1.9% for 90% and 100% share of virtual school respectively. Looking at the graph overall, the line of best fit seems to have a negative slop compared to the line of best fit computed for the other graphs. 


# Discussion

## Interesting point 1

## Change in Test Scores of 2022

In the near identical Replication of the graph in the original paper(@fig-InPersonAttendence), we have included the year 2022 in the change in test scores, where the original paper failed to do so. This could could be due to a variety of reasons, among them possibly the fact that the original authors may have thought it to be unimportant, or they may have been trying to make the change in test scores to seem more permanent, as if the the virus had more long term effect. While the test scores did not increase as much as they fell, they increased never the less, although we do not have data for spring 2023, it is possible we could have made further conclusions with this additional information.

## Change in enrollment rate in kindergarden vs all

The third set of graphs that were shown in the results section was graphs that showed the change in enrollment with the shares of in-person school that started (@fig-EnrollAll, @fig-EnrollKinder). As discussed in the section it seemed like the overall trend of the graph was that enrollment rates were higher the more in-person school there was. This was seen in both cases but more predominantly in the kindergarten graph (@fig-EnrollKinder). This result was interesting as this shows with younger ages the enrollment rates being higher means the children enjoy going to school in person. This is important as kindergarten is when most children start developing the fundamental social skills required to interact with other people.

## Ethics and Bias
  Some of the ethical concerns of this research is the idea of consent. All of this data is publicaly available so, it is more important that consent is gathered when performing these surveys. Other ethical concerns lie in the area of having the choice of in-person vs virtual schooling. Although this paper shows that in-person schooling has allowed a better enrollment rate and such, however that does not necessarily mean that students were doing better in-person. many studies during the pandemic found that course average had been either very high due to the fact that classes were taken online and students had the flexibility required to maintain a proper school-life balance but on the other hand other districts saw lower averages as the district school board would decide to test on harder material to fight against the real possibility of plagiarism during virtual school. Plagiarism is another big ethics topic to cover when talking about virtual school as attending a school online makes it very difficult for educators to monitor every students move and take action again every academic offence case that comes up. This leads to a difficult question for policy makers on whether they allow the virtual school format with stricter rules to combat such offence or to come up with a better more relaxed approach where students aren't restricted by the fear of accidentally commiting plagiarism. 
  
  In terms of bias in this paper, there is a fair amount. Most of the data would not be considered bias as the data was obtained from a reputable government website which by law requires them to publish truthful data. Also the fact that the US is considered a developped country also diminishes the chance of bias in missing data from educational districts, as they are all accounted for. A possible bias could be the fact that the data only looks at the enrollment rates for grade 12 and lower where in the states this is almost mandatory. This means that change in enrollment is never due to the fact that students just stopped going to school. A possible way to combat this bias is to have the same data but for university/college students or for the workplace. This allows a more indepth analysis about the mental health issues that might surround people during the pandemic, resulting in them either having higher productivity -which would be showcased by an increase in "enrollment" as there is more virtual options- or a decrease in productivity resulting in dropping out - which would be showcased in the graph as either an increase in "enrollment" as more in-person options open up pr decrease in "enrollment" as more virtual option open up. 



## weakness and limitations
Do our own part (how to solve it!)


## Furthur research 
As stated previously we do not have data for the test scores in 2023, which could give us additional information about how the pandemic affected students grades, as well as the younger students who were in elementary and middle schools when the pandemic began and how their educational foundations may have been affected.

## how to solve the limitations
Furthermore, the best way to see the how test scores would change would be to wait several years and analyze the data that would be collected after a few ages complete these tests. While in the short term the test scores of Spring 2023 would be benefitial information to have before making any more conclusions than we have.

## Furthur questions?

# Appendix

```{r}
#| echo: false
#| warning: false
#| label: tbl-InperAll
#| tbl-cap: Enrollment rates throughout all districts in the U.S. 2020 catagorized by share of inperson learning
weighted_means_all |> kable(format = "pipe", padding = 2, col.names = c("Share of inperson", "average total enrollment rate"))
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-KinderAll
#| tbl-cap: Enrollment rates for kindergartners throughout all districts in the U.S. 2020 catagorized by share of inperson learning
weighted_means_kinder |> kable(format = "pipe", padding = 2, col.names = c("Share of inperson", "average kindergarten enrollment rate"))
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-InperAll19
#| tbl-cap: Enrollment rates throughout all districts in the U.S. 2019 catagorized by share of inperson learning
mean_enroll_2019_all |> kable(format = "pipe", padding = 2, col.names = c("Share of inperson", "average enrollment rate"))
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-VirtAll
#| tbl-cap: Enrollment rates throughout all districts in the U.S. 2020 catagorized by share of virtual learning
virtual_2020_mean |> kable(format = "pipe", padding = 2, col.names = c("Share of virtual", "average enrollment rate"))
```


# Reference
