---
title: "COVIDSchoolClosures"
author: "Christina, Yan, Sakura"
format: pdf
---
```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

library(knitr)
library(janitor)
library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(haven)
library(tsibble)
library(readr)
library(dplyr)
library(ggplot2)

```

# Introduction


# Data

## Data Source and Collection //Methodology
 Basically cite
 
## Data Cleaning
What data set did we clean and why. Explain the variable here too ig



# Results
The graphs we made and describe the trends. Only talk abotu results not what they mean



## First Graph




##Second Graph



## Third Graph
```{r}
## Only run once cause super big data which will slow down
raw_d <- read_dta("../../inputs/data/nces_district_enrollment_2018_2020.dta")
other_raw <- read_csv("../../inputs/data/District_Overall_Shares.csv")
other_raw <- other_raw|>
  rename(leaid = "NCESDistrictID")
other_raw$leaid <- as.character(other_raw$leaid)
```





```{r}

data <- raw_d |>
  filter(grade == 0 | grade == 99) |>
  filter(race == 99)


data <- subset(data, select = -race)

# Drop observations where enrollment is less than 0
data <- data[data$enrollment >= 0, ]

# Drop observations where fips is greater than 56
data <- data[data$fips <= 56, ]

data <- data %>%
  filter(!is.na(sex))


# Reshape 'enrollment' wide by 'leaid', 'year', and 'grade'
data_wide <- pivot_wider(data,
                            names_from = sex,
                            values_from = enrollment) |>
  rename(
    enrollment1 = `1`,
    enrollment2 = `2`,
    enrollment9 = `9`,
    enrollment99 = `99`)

# Drop the 'enrollment9' variable
data_wide$enrollment9 <- NULL

# Drop observations where 'enrollment99' is missing or less than or equal to 0
data_wide <- data_wide[!is.na(data_wide$enrollment99) & data_wide$enrollment99 > 0, ]

# Reshape 'enrollment1', 'enrollment2', and 'enrollment99' wide by 'leaid' and 'year'
data_final <- pivot_wider(data_wide,
                            names_from = grade,
                            values_from = c(enrollment1, enrollment2, enrollment99))
# Can get rid of fips

data_final <- subset(data_final, select = -fips)

# Rename columns
colnames(data_final) <- c("year", "leaid", 
                              "enroll_male_kinder", "enroll_female_kinder", 
                              "enroll_male_total", "enroll_female_total", 
                              "enroll_all_kinder", "enroll_all_total")

# Convert 'leaid' to numeric (assuming it's a string)
data$leaid <- as.numeric(data$leaid)
other_raw$leaid <- as.numeric(other_raw$leaid)



# Link data with another dataset (replace 'other_data' with your actual dataset)
# Assuming 'leaid' is the common identifier
merged_data <- merge(data_final, other_raw, by = "leaid")


# Get additional variables (share_inperson, share_hybrid, share_virtual)
# from the linked dataset 
selected_vars <- linked_data[, c("share_inperson", "share_hybrid", "share_virtual")]

# Assuming 'year' is numeric (e.g., 2019, 2020)
merged_data$date <- as.Date(paste0(merged_data$year, "-01-01"))


my_tsibble <- as_tsibble(merged_data, key = leaid, index = date)


# Assuming you have a tsibble called my_tsibble with appropriate columns

# Calculate change_kinder
my_tsibble <- my_tsibble %>%
  mutate(change_kinder = enroll_all_kinder - lag(enroll_all_kinder),
         perc_change_kinder = change_kinder / lag(enroll_all_kinder))

# Calculate change_all
my_tsibble <- my_tsibble %>%
  mutate(change_all = enroll_all_total - lag(enroll_all_total),
         perc_change_all = change_all / lag(enroll_all_total))

# Calculate change_male
my_tsibble <- my_tsibble %>%
  mutate(change_male = enroll_male_kinder - lag(enroll_male_kinder),
         perc_change_male = change_male / lag(enroll_male_kinder))

# Calculate change_female
my_tsibble <- my_tsibble %>%
  mutate(change_female = enroll_female_kinder - lag(enroll_female_kinder),
         perc_change_female = change_female / lag(enroll_female_kinder))


# Drop rows where absolute change_all is greater than enroll_all_total
my_tsibble <- my_tsibble %>%
  filter(abs(change_all) <= enroll_all_total)

# Drop rows where absolute change_kinder is greater than enroll_all_total
my_tsibble <- my_tsibble %>%
  filter(abs(change_kinder) <= enroll_all_total)

my_tsibble <- my_tsibble %>%
  mutate(inperson_bin = cut(share_inperson, breaks = 11))

# Compute weighted mean change in enrollment for each bin
weighted_means_all <- my_tsibble %>%
  group_by(inperson_bin) %>%
  summarise(
    weighted_mean_change = weighted.mean(perc_change_all, enroll_all_total)
  )

weighted_means_kinder <- my_tsibble %>%
  group_by(inperson_bin) %>%
  summarise(
    weighted_mean_change = weighted.mean(perc_change_kinder, enroll_all_kinder)
  )


##Graph
ggplot(weighted_means, aes(x = inperson_bin , y = weighted_mean_change)) +
  geom_point() +
  labs(
    title = "Percent Change in Overall District Enrollment, Fall 2019-Fall 2020",
    x = "Share of School Year with In-Person Learning Offered, 2020-2021",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal()

ggplot(weighted_means_kinder, aes(x = inperson_bin, y = weighted_mean_change)) +
  geom_point() +
  labs(
    title = "Percent Change in Kindergarten Enrollment, Fall 2019-Fall 2020",
    x = "Share of School Year with In-Person Learning Offered, 2020-2021",
    y = "Percent Change in Enrollment"
  ) +
  theme_minimal()


```



# Discussion

## Interesting point 1


##Intreesting point 2

## interesting point 3



##Ethics and Bias
could talk about mental health maybe but it might apply to other "interesting point"

## weakness and limitations


## how to solve the limitations


## Furthur questions?


# Appendix



# Reference

